---
title: "Composition & Functional Overview"
output: html_document
date: "2024-01-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r libraries, comment="", echo=FALSE, message=FALSE, warning=FALSE}
library(hilldiv2) 
library(janitor)
library(ggh4x)
library(gridExtra)
library(ggpubr)
library(tidyverse)
library(ggplot2)
library(vegan)
library(dplyr)
library(tidyr)
library(distillR)

```


# Taxonomic composition

```{r loaddata, comment="", echo=FALSE, message=FALSE, warning=FALSE}
load("resources/data/data.Rdata")
```

```{r color, comment="", echo=FALSE, message=FALSE, warning=FALSE}
#Get phylum colors from the EHI standard

genome_metadata <- genome_metadata %>%
  mutate_at(vars(phylum), ~ str_replace(., "[dpcofgs]__", ""))

phylum_colors <- read_tsv("https://raw.githubusercontent.com/earthhologenome/EHI_taxonomy_colour/main/ehi_phylum_colors.tsv") %>%
  mutate_at(vars(phylum), ~ str_replace(., "[dpcofgs]__", ""))%>%
  right_join(genome_metadata, by=join_by(phylum == phylum)) %>%
    arrange(match(genome, tree$tip.label)) %>%
    dplyr::select(phylum, colors) %>%
    unique() %>%
    arrange(phylum) %>%
    dplyr::select(colors) %>%
    pull()
```

```{r barplot1, comment="", echo=FALSE, message=FALSE, warning=FALSE}
genome_counts %>%
  mutate_at(vars(-genome),~./sum(.)) %>% #apply TSS nornalisation
  pivot_longer(-genome, names_to = "sample", values_to = "count") %>% #reduce to minimum number of columns
  left_join(., genome_metadata, by = join_by(genome == genome)) %>% #append genome metadata
  left_join(., sample_metadata, by = join_by(sample == sample)) %>% #append sample metadata
  ggplot(., aes(x=sample,y=count, fill=phylum, group=phylum)) + #grouping enables keeping the same sorting of taxonomic units
    geom_bar(stat="identity", colour="white", linewidth=0.1) + #plot stacked bars with white borders
    scale_fill_manual(values=phylum_colors) +
    labs(y = "Relative abundance") +
  facet_grid(.~region,  scales="free_x")+
#    facet_nested(.~Individual+Sample_type,  scales="free_x") +
    guides(fill = guide_legend(ncol = 1)) +
    theme(axis.text.x=element_blank(),
          axis.title.x = element_blank(),
          axis.ticks.x=element_blank(),
          panel.background = element_blank(),
          panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(linewidth = 0.5, linetype = "solid", colour = "black")) +
   labs(fill="Phylum")
```

# Alpha diversity
## Alpha diversity means by sample type
```{r alpha_div, comment="", echo=FALSE, message=FALSE, warning=FALSE}
#Calculate Hill numbers
richness <- genome_counts %>% 
            column_to_rownames(var="genome") %>% 
            dplyr::select(where(~!all(. == 0))) %>% 
            hilldiv(.,q=0) %>% 
            t() %>% 
            as.data.frame() %>%
            dplyr::rename(richness=1) %>%
            rownames_to_column(var="sample")

neutral <- genome_counts %>% 
            column_to_rownames(var="genome") %>% 
            dplyr::select(where(~!all(. == 0))) %>% 
            hilldiv(.,q=1) %>% 
            t() %>% 
            as.data.frame() %>%
            dplyr::rename(neutral=1) %>%
            rownames_to_column(var="sample")

phylogenetic <- genome_counts %>% 
            column_to_rownames(var="genome") %>% 
            dplyr::select(where(~!all(. == 0))) %>% 
            hilldiv(.,q=1,tree=tree) %>% 
            t() %>% 
            as.data.frame() %>%
            dplyr::rename(phylogenetic=1) %>%
            rownames_to_column(var="sample")

# Aggregate basal GIFT into elements
dist <- genome_gifts %>%
    to.elements(., GIFT_db) %>%
    traits2dist(., method="gower")

genome_counts_filt <- genome_counts[genome_counts$genome %in% rownames(genome_gifts),] 
rownames(genome_counts_filt) <- NULL
functional <- genome_counts_filt %>% 
            column_to_rownames(var="genome") %>% 
            dplyr::select(where(~!all(. == 0))) %>% 
            hilldiv(.,q=1,dist=dist) %>% 
            t() %>% 
            as.data.frame() %>%
            dplyr::rename(functional=1) %>%
            rownames_to_column(var="sample") %>%
            mutate(functional = if_else(is.nan(functional), 1, functional))

# Merge all metrics
alpha_div <- richness %>%
      full_join(neutral,by=join_by(sample==sample)) %>%
      full_join(phylogenetic,by=join_by(sample==sample)) %>%
      full_join(functional,by=join_by(sample==sample)) %>%
      left_join(., sample_metadata, by = join_by(sample == sample))


richness_mean <- alpha_div %>%
  group_by(region) %>%
  dplyr::summarise_at(.vars = names(.)[2], .funs = c("Richness mean" = "mean", "Richness sd" = "sd")) %>%
  dplyr::rename("Region"="region")

neutral_mean <- alpha_div %>%
  group_by(region) %>%
  dplyr::summarise_at(.vars = names(.)[3], .funs = c("Neutral mean" = "mean", "Neutral sd" = "sd"))

phylogenetic_mean <- alpha_div %>%
  group_by(region) %>%
  dplyr::summarise_at(.vars = names(.)[4], .funs = c("Phylogenetic mean" = "mean", "Phylogenetic sd" = "sd"))

functional_mean <- alpha_div %>%
  group_by(region) %>%
  dplyr::summarise_at(.vars = names(.)[5], .funs = c("Functional mean" = "mean", "Functional sd" = "sd"))

cbind(richness_mean, neutral_mean[, 2:3], phylogenetic_mean[, 2:3], functional_mean[, 2:3]) %>% #
  as.data.frame()%>%
  t() %>%
  row_to_names(row_number = 1) %>%
  as.data.frame() %>%
  mutate_if(is.character, as.numeric)  %>%
  knitr::kable(.,digits = c(3,3))
```

## Plot diversities
```{r alpha_div_plot1, comment="", echo=FALSE, message=FALSE, warning=FALSE}
alpha_div_pivot <- richness %>%
  full_join(neutral,by=join_by(sample==sample)) %>%
  full_join(phylogenetic,by=join_by(sample==sample)) %>%
  full_join(functional,by=join_by(sample==sample)) %>%
  pivot_longer(-sample, names_to = "data", values_to = "value") %>%
  left_join(., sample_metadata, by = join_by(sample == sample))

alpha_div_pivot %>%
  ggplot(aes(x=value, y=sample)) +
  geom_bar(stat='identity', fill="#6c9ebc") +
  facet_nested(region ~ data,  scales="free")+
  coord_cartesian(xlim = c(1, NA)) +
  theme_classic() +
  theme(strip.background = element_blank(),
    panel.grid.minor.x = element_line( size=.1, color="grey" ),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1))

```


```{r alpha_div_rich, comment="", echo=FALSE, message=FALSE, warning=FALSE}
alpha_colors <- c("#e5bd5b", "#6b7398", "#76b183", "#d57d2c", "#2a2d26", "#f9d4cc", "#3c634e", "#ea68c3")
group_n <- alpha_div %>%
  select(region) %>%
  pull() %>%
  unique() %>%
  length()
plot1 <- alpha_div %>%
  ggplot(aes(x = region, y = richness, group = region, color = region, fill = region)) +
  geom_jitter(width = 0.05, size = 1.5, show.legend = FALSE) +
  geom_violin(alpha = 0.2, width = 0.3, show.legend = FALSE) +
  scale_color_manual(values = alpha_colors[c(1:group_n)]) +
  scale_fill_manual(values = paste0(alpha_colors[c(1:group_n)], "50")) +
  stat_compare_means(show.legend = FALSE) +
  theme(axis.text.x = element_text(vjust = 0.5),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(face = "bold", size = 12),
    axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
    axis.title.x = element_text(margin = margin(t = 20, r = 0, b = 0, l = 0)),
    panel.background = element_blank(),
    # panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(size = 0.5, linetype = "solid", colour = "black")) +
  guides(colour = guide_legend(override.aes = list(size = 6))) +
  labs(x = "Region", y = "Richness")
```

```{r alpha_div_neutral, comment="", echo=FALSE, message=FALSE, warning=FALSE}
plot2 <- alpha_div%>%
  ggplot(aes(x = region, y = neutral, group = region, color = region, fill = region)) +
  geom_jitter(width = 0.05, size = 1.5, show.legend = FALSE) +
  geom_violin(alpha = 0.2, width = 0.3, show.legend = FALSE) +
  scale_color_manual(values = alpha_colors[c(1:group_n)]) +
  scale_fill_manual(values = paste0(alpha_colors[c(1:group_n)], "50")) +
  stat_compare_means(show.legend = FALSE) +
  theme(axis.text.x = element_text(vjust = 0.5),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(face = "bold", size = 12),
    axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
    axis.title.x = element_text(margin = margin(t = 20, r = 0, b = 0, l = 0)),
    panel.background = element_blank(),
    # panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(size = 0.5, linetype = "solid", colour = "black")) +
  guides(colour = guide_legend(override.aes = list(size = 6)))+
  labs(x = "Region", y = "Neutral")
```

```{r alpha_div_phylo, comment="", echo=FALSE, message=FALSE, warning=FALSE}
plot3 <- alpha_div%>%
  ggplot(aes(x = region, y = phylogenetic, group = region, color = region, fill = region)) +
  geom_jitter(width = 0.05, size = 1.5, show.legend = FALSE) +
  geom_violin(alpha = 0.2, width = 0.3, show.legend = FALSE) +
  scale_color_manual(values = alpha_colors[c(1:group_n)]) +
  scale_fill_manual(values = paste0(alpha_colors[c(1:group_n)], "50")) +
  stat_compare_means(show.legend = FALSE) +
  theme(axis.text.x = element_text(vjust = 0.5),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(face = "bold", size = 12),
    axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
    axis.title.x = element_text(margin = margin(t = 20, r = 0, b = 0, l = 0)),
    panel.background = element_blank(),
    # panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(size = 0.5, linetype = "solid", colour = "black")) +
  guides(colour = guide_legend(override.aes = list(size = 6)))+
  labs(x = "Region", y = "Phylogenetic")
```
```{r alpha_div_func, comment="", echo=FALSE, message=FALSE, warning=FALSE}
plot4 <- alpha_div%>%
  ggplot(aes(x = region, y = functional, group = region, color = region, fill = region)) +
  geom_jitter(width = 0.05, size = 1.5, show.legend = FALSE) +
  geom_violin(alpha = 0.2, width = 0.3, show.legend = FALSE) +
  scale_color_manual(values = alpha_colors[c(1:group_n)]) +
  scale_fill_manual(values = paste0(alpha_colors[c(1:group_n)], "50")) +
  stat_compare_means(show.legend = FALSE) +
  theme(axis.text.x = element_text(vjust = 0.5),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(face = "bold", size = 12),
    axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
    axis.title.x = element_text(margin = margin(t = 20, r = 0, b = 0, l = 0)),
    panel.background = element_blank(),
    # panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(size = 0.5, linetype = "solid", colour = "black")) +
  guides(colour = guide_legend(override.aes = list(size = 6)))+
  labs(x = "Region", y = "Functional")
```

```{r div_plot_together, comment="", echo=FALSE, message=FALSE, warning=FALSE, fig.height=8, fig.width=10, fig.fullwidth=TRUE}
grid.arrange(arrangeGrob(plot1,plot2,plot3,plot4, ncol = 2)) 
```

# Beta diversity
```{r beta_div, comment="", echo=FALSE, message=FALSE, warning=FALSE, results='hide', eval=FALSE}
beta_q0 <- genome_counts%>%
  column_to_rownames(., "genome")%>%
  hillpair(., q = 0)

beta_q1n <- genome_counts%>%
  column_to_rownames(., "genome")%>%
  hillpair(., q = 1)

beta_q1p <- genome_counts%>%
  column_to_rownames(., "genome")%>%
  hillpair(., q = 1, tree = tree)

beta_div_func <- genome_counts_filt%>%
  column_to_rownames(., "genome")%>%
  hillpair(.,q=1,dist=dist)

save(beta_q0,beta_q1n, beta_q1p, beta_div_func, file = "resources/data/beta_div.Rdata")
```

```{r beta_div_open, comment="", echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
load("resources/data/beta_div.Rdata")
```

```{r betan, comment="", echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
beta_metric <- beta_q1n$S

beta_q1n_nmds <- beta_metric %>%
  vegan::metaMDS(., trymax = 500, k = 2, verbosity = FALSE) %>%
  vegan::scores() %>%
  as_tibble(., rownames = "sample") %>%
  dplyr::left_join(sample_metadata, by = join_by(sample == sample))

group_n <- length(unique(beta_q1n_nmds$region))
beta_colors <- c("#e5bd5b", "#6b7398", "#76b183", "#d57d2c", "#2a2d26", "#f9d4cc", "#3c634e", "#ea68c3")
```

## Neutral beta diversity
```{r beta_div_neutral, comment="", echo=FALSE, message=FALSE, warning=FALSE}

# pdf(paste(workingdir,"/x.pdf",sep=""),width=14, height=9)
q1n_nmds <- beta_q1n_nmds %>%
  group_by(region) %>%
  mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
  ungroup()
ggplot(q1n_nmds, aes(x = NMDS1, y = NMDS2, color = region)) +
  scale_color_manual(values = beta_colors[c(1:group_n)]) +
  scale_shape_manual(values = 1:10) +
  geom_point(size = 4) +
  #   stat_ellipse(aes(color = beta_q1n_nmds$Groups))+
  geom_segment(aes(x = x_cen, y = y_cen, xend = NMDS1, yend = NMDS2), alpha = 0.9) +
  theme_classic() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 20, face = "bold"),
    axis.text = element_text(face = "bold", size = 18),
    panel.background = element_blank(),
    axis.line = element_line(size = 0.5, linetype = "solid", colour = "black"),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18),
    legend.position = "right", legend.box = "vertical"
  )
#+ geom_text(aes(label = Sample), size=8)
# dev.off()
```

***Homogeneity of variance***

```{r permu1, comment="", echo=FALSE, message=FALSE, warning=FALSE}
ps.disper.neutral <- betadisper(beta_metric, sample_metadata$region) 
permutest(ps.disper.neutral, pairwise = TRUE) 
```

***Permanova***
```{r adonisbeta2, comment="", echo=FALSE, message=FALSE, warning=FALSE}
#### Adonis
metarow <- column_to_rownames(sample_metadata, "sample")

adonis2(beta_metric ~ region+sex, data = metarow[labels(beta_metric), ], permutations = 999) %>%
  as.matrix() %>%
  knitr::kable(., digits = c(0,3,3,3,3))
```

```{r betap, comment="", echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
beta_p_metric <- beta_q1p$S

beta_q1p_nmds <- beta_p_metric %>%
  vegan::metaMDS(., trymax = 500, k = 2, verbosity = FALSE) %>%
  vegan::scores() %>%
  as_tibble(., rownames = "sample") %>%
  dplyr::left_join(sample_metadata, by = join_by(sample == sample))

group_n <- length(unique(beta_q1p_nmds$region))
beta_colors <- c("#e5bd5b", "#6b7398", "#76b183", "#d57d2c", "#2a2d26", "#f9d4cc", "#3c634e", "#ea68c3")
```

## Phylogenetic beta diversity
```{r  beta_div_phylo, comment="", echo=FALSE, message=FALSE, warning=FALSE}
# pdf(paste(workingdir,"/betaB692_caecum.pdf",sep=""),width=14, height=9)
q1p_nmds <- beta_q1p_nmds %>%
  group_by(region) %>%
  mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
  ungroup()
ggplot(q1p_nmds, aes(x = NMDS1, y = NMDS2, color = region)) +
  scale_color_manual(values = beta_colors[c(1:group_n)]) +
  scale_shape_manual(values = 1:10) +
  geom_point(size = 4)  +
  geom_segment(aes(x = x_cen, y = y_cen, xend = NMDS1, yend = NMDS2), alpha = 0.9) +
  theme_classic() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 20, face = "bold"),
    axis.text = element_text(face = "bold", size = 18),
    panel.background = element_blank(),
    axis.line = element_line(size = 0.5, linetype = "solid", colour = "black"),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18),
    legend.position = "right", legend.box = "vertical"
  )
#+ geom_text(aes(label = Sample), size=8)
# dev.off()
```

***Homogeneity of variance***

```{r permu2, comment="", echo=FALSE, message=FALSE, warning=FALSE}
ps.disper.phylo <- betadisper(beta_p_metric, sample_metadata$region) 
permutest(ps.disper.phylo, pairwise = TRUE) 
```

***Permanova***
```{r adonisbeta3, comment="", echo=FALSE, message=FALSE, warning=FALSE}
#### Adonis
metarow <- column_to_rownames(sample_metadata, "sample")

adonis2(beta_p_metric ~ region+sex, data = metarow[labels(beta_p_metric), ], permutations = 999) %>%
  as.matrix() %>%
  knitr::kable(., digits = c(0,3,3,3,3))
```

```{r betaf, comment="", echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
beta_f_metric <- beta_div_func$S

beta_f_nmds <- beta_f_metric %>%
  vegan::metaMDS(., trymax = 500, k = 2, verbosity = FALSE) %>%
  vegan::scores() %>%
  as_tibble(., rownames = "sample") %>%
  dplyr::left_join(sample_metadata, by = join_by(sample == sample))

group_n <- length(unique(beta_f_nmds$region))
beta_colors <- c("#e5bd5b", "#6b7398", "#76b183", "#d57d2c", "#2a2d26", "#f9d4cc", "#3c634e", "#ea68c3")
```

## Functional beta diversity
```{r  beta_div_func, comment="", echo=FALSE, message=FALSE, warning=FALSE}

# pdf(paste(workingdir,"/x.pdf",sep=""),width=14, height=9)
f_nmds <- beta_f_nmds %>%
  group_by(region) %>%
  mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
  ungroup()
ggplot(f_nmds, aes(x = NMDS1, y = NMDS2, color = region)) +
  scale_color_manual(values = beta_colors[c(1:group_n)]) +
  scale_shape_manual(values = 1:10) +
  geom_point(size = 4) +
  #   stat_ellipse(aes(color = beta_q1n_nmds$Groups))+
  geom_segment(aes(x = x_cen, y = y_cen, xend = NMDS1, yend = NMDS2), alpha = 0.9) +
  theme_classic() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 20, face = "bold"),
    axis.text = element_text(face = "bold", size = 18),
    panel.background = element_blank(),
    axis.line = element_line(size = 0.5, linetype = "solid", colour = "black"),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18),
    legend.position = "right", legend.box = "vertical"
  )
#+ geom_text(aes(label = Sample), size=8)
# dev.off()
```

***Homogeneity of variance***
```{r permu3, comment="", echo=FALSE, message=FALSE, warning=FALSE}
ps.disper.functional <- betadisper(beta_f_metric, sample_metadata$region) 
permutest(ps.disper.functional, pairwise = TRUE) 
```

***Permanova***
```{r adonisbeta4, comment="", echo=FALSE, message=FALSE, warning=FALSE}
#### Adonis

metarow <- column_to_rownames(sample_metadata, "sample")

adonis2(beta_f_metric ~ region, data = metarow[labels(beta_f_metric), ], permutations = 999) %>%
  as.matrix() %>%
  knitr::kable(., digits = c(0,3,3,3,3))
```

```{r gift, comment="", echo=FALSE, message=FALSE, warning=FALSE}
#Aggregate bundle-level GIFTs into the compound level
GIFTs_elements <- to.elements(genome_gifts,GIFT_db)
GIFTs_elements_filtered <- GIFTs_elements[rownames(GIFTs_elements) %in% genome_counts_filt$genome,]
GIFTs_elements_filtered <- as.data.frame(GIFTs_elements_filtered) %>% 
  select_if(~ !is.numeric(.) || sum(.) != 0)

#Aggregate element-level GIFTs into the function level
GIFTs_functions <- to.functions(GIFTs_elements_filtered,GIFT_db)

#Aggregate function-level GIFTs into overall Biosynthesis, Degradation and Structural GIFTs
GIFTs_domains <- to.domains(GIFTs_functions,GIFT_db)

#Get community-weighed average GIFTs per sample
genome_counts_row <- genome_counts %>%
  mutate_at(vars(-genome),~./sum(.)) %>% 
  column_to_rownames(., "genome") 

GIFTs_elements_community <- to.community(GIFTs_elements_filtered,genome_counts_row,GIFT_db)
GIFTs_functions_community <- to.community(GIFTs_functions,genome_counts_row,GIFT_db)
GIFTs_domains_community <- to.community(GIFTs_domains,genome_counts_row,GIFT_db)
```

### Community elements differences
```{r comunity_elem, comment="", echo=FALSE, message=FALSE, warning=FALSE}
element_gift <- GIFTs_elements_community %>% 
  as.data.frame() %>% 
  rownames_to_column(., "sample") %>% 
  merge(., sample_metadata[c(1,4)], by="sample")
```

```{r commun_wilcox_elem, comment="", echo=FALSE, message=FALSE, warning=FALSE}
uniqueGIFT_db<- unique(GIFT_db[c(2,4,5,6)]) %>% unite("Function",Function:Element, sep= "_", remove=FALSE)

significant_elements <- element_gift %>%
    pivot_longer(-c(sample,region), names_to = "trait", values_to = "value") %>%
    group_by(trait) %>%
    summarise(p_value = wilcox.test(value ~ region)$p.value) %>%
    mutate(p_adjust=p.adjust(p_value, method="BH")) %>%
    filter(p_adjust < 0.05)%>%
  rownames_to_column(., "Elements")  %>%
  left_join(.,uniqueGIFT_db[c(1,3)],by = join_by(trait == Code_element))

element_gift_t <- element_gift  %>% 
  t() %>%
  row_to_names(row_number = 1) %>%
  as.data.frame() %>%
  mutate_if(is.character, as.numeric)  %>%
  rownames_to_column(., "trait")

element_gift_filt <- subset(element_gift_t, trait %in% significant_elements$trait) %>% 
  t() %>%
  row_to_names(row_number = 1) %>%
  as.data.frame() %>%
  mutate_if(is.character, as.numeric)  %>%
  rownames_to_column(., "sample")%>% 
  left_join(., sample_metadata[c(1,4)], by = join_by(sample == sample))

element_gift_filt %>%
  select(-sample)%>%
  group_by(region)  %>%
  summarise(across(everything(), mean))%>%
   t() %>%
  row_to_names(row_number = 1) %>%
  as.data.frame() %>%
  mutate_if(is.character, as.numeric)  %>%
  rownames_to_column(., "Elements")  %>%
  left_join(.,uniqueGIFT_db[c(1,3)],by = join_by(Elements == Code_element))
```

```{r elements_plot, comment="", echo=FALSE, message=FALSE, warning=FALSE}
element_gift_names <- element_gift_filt%>%
  select(-region)%>%
   t() %>%
  row_to_names(row_number = 1) %>%
  as.data.frame() %>%
  mutate_if(is.character, as.numeric)  %>%
  rownames_to_column(., "Elements")  %>%
  left_join(.,uniqueGIFT_db[c(1,3)],by = join_by(Elements == Code_element))%>%
  select(-Elements)%>%
  select(Function, everything())%>%
   t()%>%
  row_to_names(row_number = 1) %>%
  as.data.frame() %>%
  mutate_if(is.character, as.numeric)  %>%
  rownames_to_column(., "sample")%>% 
  left_join(., sample_metadata[c(1,4)], by = join_by(sample == sample))


colNames <- names(element_gift_names)[2:87]
for(i in colNames){
  plt <- ggplot(element_gift_names, aes(x=region, y=.data[[i]], color = region)) +
    geom_boxplot(alpha = 0.2, outlier.shape = NA, width = 0.3, show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = TRUE) +
  theme_minimal() +
  theme(
    axis.line = element_line(size = 0.5, linetype = "solid", colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank())
print(plt)
}
```


### Community functions differences

```{r comunity_func, comment="", echo=FALSE, message=FALSE, warning=FALSE}
function_gift <- GIFTs_functions_community %>% 
  as.data.frame() %>% 
  rownames_to_column(., "sample") %>% 
  merge(., sample_metadata[c(1,4)], by="sample")
```

```{r commun_wilcox_func, comment="", echo=FALSE, message=FALSE, warning=FALSE}
unique_funct_db<- GIFT_db[c(3,4,5)] %>% 
  distinct(Code_function, .keep_all = TRUE)


significant_functional <- function_gift %>%
    pivot_longer(-c(sample,region), names_to = "trait", values_to = "value") %>%
    group_by(trait) %>%
    summarise(p_value = wilcox.test(value ~ region)$p.value) %>%
    mutate(p_adjust=p.adjust(p_value, method="BH")) %>%
    filter(p_adjust < 0.05)%>%
  left_join(.,unique_funct_db[c(1,3)],by = join_by(trait == Code_function))

```
```{r  function_sig, comment="", echo=FALSE, message=FALSE, warning=FALSE}
function_gift_t <- function_gift  %>% 
  select(-region)  %>% 
  t() %>%
  row_to_names(row_number = 1) %>%
  as.data.frame() %>%
  mutate_if(is.character, as.numeric)  %>%
  rownames_to_column(., "trait")

function_gift_filt <- subset(function_gift_t, trait %in% significant_functional$trait) %>% 
  t() %>%
  row_to_names(row_number = 1) %>%
  as.data.frame() %>%
  mutate_if(is.character, as.numeric)  %>%
  rownames_to_column(., "sample")%>% 
  left_join(., sample_metadata[c(1,4)], by = join_by(sample == sample))


function_gift_filt %>%
  select(-sample)%>%
  group_by(region)  %>%
  summarise(across(everything(), mean))%>%
   t() %>%
  row_to_names(row_number = 1) %>%
  as.data.frame() %>%
  mutate_if(is.character, as.numeric)  %>%
  rownames_to_column(., "Code_function")  %>%
  left_join(.,unique_funct_db[c(1,3)],by = join_by(Code_function == Code_function))
```
```{r function_plot, comment="", echo=FALSE, message=FALSE, warning=FALSE}
function_gift_names <- function_gift_filt%>%
  select(-region)%>%
   t() %>%
  row_to_names(row_number = 1) %>%
  as.data.frame() %>%
  mutate_if(is.character, as.numeric)  %>%
  rownames_to_column(., "Code_function")  %>%
  left_join(.,unique_funct_db[c(1,3)],by = join_by(Code_function == Code_function))%>%
  select(-Code_function)%>%
  select(Function, everything())%>%
   t()%>%
  row_to_names(row_number = 1) %>%
  as.data.frame() %>%
  mutate_if(is.character, as.numeric)  %>%
  rownames_to_column(., "sample")%>% 
  left_join(., sample_metadata[c(1,4)], by = join_by(sample == sample))


colNames <- names(function_gift_names)[2:13]
for(i in colNames){
  plt <- ggplot(function_gift_names, aes(x=region, y=.data[[i]], color = region)) +
    geom_boxplot(alpha = 0.2, outlier.shape = NA, width = 0.3, show.legend = FALSE) +
  geom_jitter(width = 0.1, show.legend = TRUE) +
  theme_minimal() +
  theme(
    axis.line = element_line(size = 0.5, linetype = "solid", colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank())
print(plt)
}
```

### Community domains differences
***No differences***
```{r comunity_dom, comment="", echo=FALSE, message=FALSE, warning=FALSE}
domain_gift <- GIFTs_domains_community %>% 
  as.data.frame() %>% 
  rownames_to_column(., "sample") %>% 
  merge(., sample_metadata[c(1,4)], by="sample")
```

```{r commun_wilcox_dom, comment="", echo=FALSE, message=FALSE, warning=FALSE}
unique_domain_db<- GIFT_db[c(4)] %>% 
  distinct(Domain, .keep_all = TRUE)

significant_domain <- domain_gift %>%
    pivot_longer(-c(sample,region), names_to = "trait", values_to = "value") %>%
    group_by(trait) %>%
    summarise(p_value = wilcox.test(value ~ region, exact = FALSE, alternative = "less")$p.value) %>%
    mutate(p_adjust=p.adjust(p_value, method="BH"))
```

