---
title: "Different_abundance"
output: html_document
date: "2023-11-15"
---
```{r clr_transf,comment="", echo=FALSE}
physeq_clr <- microbiome::transform(physeq_all, 'clr')
physeq_phylum_clr <- microbiome::aggregate_taxa(physeq_clr, 'phylum')
```

### Differences in phyla between the two groups of sled dogs

```{r phylum3, comment="", echo=FALSE, eval=FALSE}
#### Ittoqqortoormii
physeq_phylum_clr <- microbiome::aggregate_taxa(physeq_clr, 'phylum')
physeq_phylum_rel1 <- subset_samples(physeq_phylum_clr, region == "Ittoqqortoormii")
physeq_phylum_rel1 <- prune_taxa(taxa_sums(physeq_phylum_rel1)>0, physeq_phylum_rel1)
table.rel1 <- physeq_phylum_rel1@otu_table
means.table.rel1 <- as.data.frame(rowMeans(table.rel1))
sd.table.rel1 <- as.data.frame(rowSds(table.rel1, useNames = TRUE))
summary.phylum1 <- merge(means.table.rel1, sd.table.rel1, by="row.names")
colnames(summary.phylum1) <- c("Phylum","Mean", "SD")
print(summary.phylum1[order(-summary.phylum1$Mean),], row.names = FALSE)
```

```{r phylum4, comment="", echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
#### Daneborg
physeq_phylum_clr <- microbiome::aggregate_taxa(physeq_clr, 'phylum')
physeq_phylum_rel1 <- subset_samples(physeq_phylum_clr, region == "Daneborg")
physeq_phylum_rel1 <- prune_taxa(taxa_sums(physeq_phylum_rel1)>0, physeq_phylum_rel1)
table.rel1 <- physeq_phylum_rel1@otu_table
means.table.rel1 <- as.data.frame(rowMeans(table.rel1))
sd.table.rel1 <- as.data.frame(rowSds(table.rel1, useNames = TRUE))
summary.phylum1 <- merge(means.table.rel1, sd.table.rel1, by="row.names")
colnames(summary.phylum1) <- c("Phylum","Mean", "SD")
print(summary.phylum1[order(-summary.phylum1$Mean),], row.names = FALSE)
```

#### Mean and sd of the phyla found in sled dogs (CLR data)

```{r mean_phyla, comment="", echo=FALSE, message=FALSE, warning=FALSE}
physeq_phylum_clr.t <- as.data.frame(t(as.matrix(physeq_phylum_clr@otu_table)))
table.W <- tibble::rownames_to_column(physeq_phylum_clr.t, "sample")
sample_type <- metadata[,c(1,4)]
table.W.meta <- merge(table.W,sample_type,by="sample")
table.W.meta <- table.W.meta[,-c(1)]
means_phylum_by_regionall <- table.W.meta %>% group_by(region) %>% summarise_at(.vars = names(.)[1:12],.funs = c(Maximum="max", Minimum="min",mean="mean", sd="sd"))
means_phylum_by_region <- table.W.meta %>% group_by(region) %>% summarise_at(.vars = names(.)[1:12],.funs = c(mean="mean", sd="sd"))
means_all <- as.data.frame(means_phylum_by_regionall)
means <- as.data.frame(means_phylum_by_region)

means <- means[,c(1,2,14,3,15,4,16,5,17,6,18,7,19,8,20,9,21,10,22,11,23,12,24,13,25)]

knitr::kable(means, format = "html", full_width = F, digits = 3) %>%
  kable_styling(latex_options="scale_down")
```

#### Phyla statistically different in the two groups (CLR data)

```{r wilcox_phyla, comment="", echo=FALSE, message=FALSE, warning=FALSE}
physeq_phylum_clr.t <- as.data.frame(t(as.matrix(physeq_all@otu_table)))
metadata <- data.frame(physeq_all@sam_data)
table.W <- tibble::rownames_to_column(physeq_phylum_clr.t, "sample")
sample_type <- metadata[,c(1,4)]
phylum_table <- merge(table.W,sample_type,by="sample")
phylum_table <- phylum_table[,-c(1)]
phylum_table.no <- phylum_table[,-ncol(phylum_table)]
all.taxa <- colnames(phylum_table.no)
Wilcox_result <- c()
for (y in all.taxa){
  res.wilcox <- wilcox.test(phylum_table[,y] ~ region, data = phylum_table,
                            exact = FALSE, alternative = "less")
  Wilcox_result <- rbind(Wilcox_result,c(res.wilcox[[1]],Pvalue=res.wilcox[[3]]))
}
rownames(Wilcox_result) <- all.taxa
Wilcox_result <- as.data.frame(Wilcox_result)

Wilcox_result$p_adjusted <- p.adjust(Wilcox_result$Pvalue, method = "fdr")

Wilcox_result_sign <- subset(Wilcox_result, p_adjusted <= 0.05) %>%
  rownames_to_column(., "Phylum")
knitr::kable(Wilcox_result_sign, format = "html", full_width = F,col.names = c("Phylum", "W", "p-value","p-value adjusted"), digits = 4) %>%
  kable_styling(latex_options="scale_down")
```

#### Mean and sd of the phyla statistically significant between the two group of sled dogs (CLR data)

```{r selec_mean_phyla, comment="", echo=FALSE, message=FALSE, warning=FALSE}
table.W_filtered <- physeq_phylum_clr.t[colnames(physeq_phylum_clr.t) %in% Wilcox_result_sign$Phylum]
sample_type <- metadata[,c(1,4)]
table.W_filtered_meta <- table.W_filtered %>%
  rownames_to_column(., "sample") %>%
  merge(.,sample_type,by="sample")
table.W_filtered_meta <- table.W_filtered_meta[,-c(1)]
means_phylum_by_region <- table.W_filtered_meta %>% group_by(region) %>% summarise_at(.vars = names(.)[1:3],.funs = c(mean="mean", sd="sd"))
means <- as.data.frame(means_phylum_by_region)
means <- means[,c(1,2,5,3,6,4,7)]
knitr::kable(means, format = "html", full_width = F, digits = 3) %>%
  kable_styling(latex_options="scale_down")
```


### Deseq
```{r load_deseq, comment="", echo=FALSE, message=FALSE, warning=FALSE}
library(DESeq2)
library(genefilter)
```

```{r deseq_spain, comment="", echo=FALSE, message=FALSE, warning=FALSE}
#physeq_caecum_aggreg <- microbiome::aggregate_taxa(physeq_caecum7, 'Genus')
#diagdds = phyloseq_to_deseq2(physeq_caecum_aggreg, ~ Treatment)

diagdds = phyloseq_to_deseq2(physeq_all, ~ Origin)

diagdds <- estimateSizeFactors(diagdds, type="poscounts",locfunc=genefilter::shorth)
diagdds = DESeq(diagdds, test="Wald", fitType="parametric")
diagdds.ins.fis <- results(diagdds, alpha=0.05, contrast=c("Origin", "Feral", "Tame"))
sigtab_diagdds.ins.fis <- diagdds.ins.fis[which(diagdds.ins.fis$padj < 0.05), ]
sigtab_diagdds_with_tax <- cbind(as(sigtab_diagdds.ins.fis, "data.frame"), as(tax_table(physeq_all)[row.names(sigtab_diagdds.ins.fis), ], "matrix"))
#sigtab_diagdds_with_tax[order(sigtab_diagdds_with_tax$baseMean, decreasing=T), ]
deseq2_group <- as.data.frame(sigtab_diagdds_with_tax)
theme_set(theme_bw())
scale_fill_discrete <- function(palname = "Set1", ...) {
  scale_fill_brewer(palette = palname, ...)
}

x = tapply(sigtab_diagdds_with_tax$log2FoldChange, sigtab_diagdds_with_tax$Order, function(x) max(x))
x = sort(x, TRUE)
sigtab_diagdds_with_tax$Family = factor(as.character(sigtab_diagdds_with_tax$Order), levels=names(x))
x = tapply(sigtab_diagdds_with_tax$log2FoldChange, sigtab_diagdds_with_tax$Genus, function(x) max(x))
x = sort(x, TRUE)
sigtab_diagdds_with_tax$Genus = factor(as.character(sigtab_diagdds_with_tax$Genus), levels=names(x))
```

```{r deseq_plot, comment="", echo=FALSE, message=FALSE, warning=FALSE}
sigtab_diagdds_with_tax %>%
ggplot(aes(x=Genus, y=log2FoldChange, color=Phylum)) + 
  geom_point(size=4) +
  geom_hline(yintercept=0) + 
  coord_flip()+
  theme(axis.text = element_text(size = 10, face="bold.italic"),
          axis.title = element_text(size = 12),
          legend.position = "bottom", 
          legend.title = element_blank())
```

#### Genera statistically different in the two groups (ANCOM-BC)
```{r}
set.seed(1234) #set seed for reproducibility
ancom_output = ancombc2(data = physeq_all, 
                  assay_name = "counts", 
                  tax_level = NULL, #change to agglomerate analysis to a higher taxonomic range
                  fix_formula = "Origin + Location", #fixed variable(s)
                  rand_formula = "(1|Location)",
                  p_adj_method = "holm", 
                  pseudo_sens = TRUE,
                  prv_cut = 0.10, 
                  lib_cut = 1000, 
                  s0_perc = 0.05,
                  group = NULL, 
                  struc_zero = FALSE, 
                  neg_lb = FALSE,
                  alpha = 0.05, 
                  n_cl = 2, 
                  verbose = TRUE,
                  global = FALSE, 
                  pairwise = FALSE, 
                  dunnet = FALSE, 
                  trend = FALSE,
                  iter_control = list(tol = 1e-5, max_iter = 20, verbose = FALSE),
                  em_control = list(tol = 1e-5, max_iter = 100),
                  lme_control = lme4::lmerControl(),
                  mdfdr_control = list(fwer_ctrl_method = "holm", B = 100), 
                  trend_control = NULL)

tax <- data.frame(physeq_all@tax_table) %>%
  rownames_to_column(., "taxon")

ancombc_table <- ancom_output$res %>%
  dplyr::select(taxon, lfc_OriginTame, p_OriginTame) %>%
  filter(p_OriginTame < 0.05) %>%
  dplyr::arrange(p_OriginTame) %>%
  merge(., tax, by="taxon")

colors_alphabetic <- ehi_phylum_colors1 %>%
  right_join(tax, by=join_by(phylum == Phylum)) %>%
  select(phylum, colors) %>%
	unique() %>%
	dplyr::arrange(phylum)

tax_table <- as.data.frame(unique(ancombc_table$Phylum))
colnames(tax_table)[1] <- "phylum"
tax_color <- merge(tax_table, colors_alphabetic, by="phylum")%>%
	dplyr::arrange(phylum) %>%
	select(colors) %>%
	pull()
```

```{r ancombc_plot, comment="", echo=FALSE, message=FALSE, warning=FALSE}
ggplot(ancombc_table, aes(x=forcats::fct_rev(taxon), y=lfc_OriginTame, color=Phylum)) + 
  geom_point(size=4) + 
  scale_color_manual(values=tax_color) + 
  geom_hline(yintercept=0) + 
  coord_flip()+
  theme(panel.background = element_blank(),
          axis.line = element_line(size = 0.5, linetype = "solid", colour = "black"),
          axis.title.x=element_blank())+
  xlab("Genus") + 
  ylab("log2FoldChange")+
  guides(col=guide_legend("Phylum"))
```



```{r ancombc2, comment="", echo=FALSE, message=FALSE, warning=FALSE}
library(ANCOMBC)
physeq.genus <- tax_glom(physeq, 'genus')
tse = mia::makeTreeSummarizedExperimentFromPhyloseq(physeq.genus)
ancombc2_out <- ancombc2(data = tse,
                         assay_name = "counts",
                         fix_formula = "region",
                         p_adj_method = "fdr",
                         prv_cut = 0,
                         group = "region",
                         struc_zero = TRUE,
                         neg_lb = TRUE,
                         # multi group comparison is deactivated automatically
                         global = TRUE)

tax <- data.frame(physeq_all@tax_table) %>%
  rownames_to_column(., "taxon")

ancombc_table <- ancom_output$res %>%
  dplyr::select(taxon, lfc_OriginTame, p_OriginTame) %>%
  filter(p_OriginTame < 0.05) %>%
  dplyr::arrange(p_OriginTame) %>%
  merge(., tax, by="taxon")

colors_alphabetic <- ehi_phylum_colors1 %>%
  right_join(taxonomyclean, by=join_by(phylum == phylum)) %>%
  select(phylum, colors) %>%
	unique() %>%
	dplyr::arrange(phylum)

tax_table <- as.data.frame(unique(ancombc_table$phylum))
colnames(tax_table)[1] <- "phylum"
tax_color <- merge(tax_table, colors_alphabetic, by="phylum")%>%
	dplyr::arrange(phylum) %>%
	select(colors) %>%
	pull()
```
```{r ancombc_plot, comment="", echo=FALSE, message=FALSE, warning=FALSE}
ggplot(ancombc_table, aes(x=forcats::fct_rev(genus), y=lfc_regionIttoqqortoormii, color=phylum)) + 
  geom_point(size=4) + 
  scale_color_manual(values=tax_color) + 
  geom_hline(yintercept=0) + 
  coord_flip()+
  xlab("Genus") + 
  ylab("log2FoldChange")+
  guides(col=guide_legend("Phylum"))
```
```{r ancombc_plot, comment="", echo=FALSE, message=FALSE, warning=FALSE}
ggplot(ancombc_table, aes(x=forcats::fct_rev(genus), y=lfc_regionIttoqqortoormii, color=phylum)) + 
  geom_point(size=4) + 
  scale_color_manual(values=tax_color) + 
  geom_hline(yintercept=0) + 
  coord_flip()+
  xlab("Genus") + 
  ylab("log2FoldChange")+
  guides(col=guide_legend("Phylum"))
```
```{r ancombc_plot1, comment="", echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
ggplot(ancombc_table, aes(x=forcats::fct_rev(reorder(genus, as.integer(factor(phylum)))), y=lfc_regionIttoqqortoormii, color=phylum)) + 
  geom_point(size=4) + 
  scale_color_manual(values=tax_color) + 
  geom_hline(yintercept=0) + 
  coord_flip()+
  xlab("Genus") + 
  ylab("log2FoldChange")+
  guides(col=guide_legend("Phylum"))
```

#### Genera statistically different in the two groups (Deseq2)

```{r deseq2, comment="", echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
#library(DESeq2)
physeq.genus <- tax_glom(physeq, 'phylum')
deseq_ob = phyloseq_to_deseq2(physeq.genus, ~ region)
#deseq_ob = phyloseq_to_deseq2(physeq, ~ region)
deseq_ob <- estimateSizeFactors(deseq_ob, type="poscounts",locfunc=genefilter::shorth)#count table is sparse with many zeroes. In that case you'd need to use a specific function first that is equipped to deal with that.
deseq_ob = DESeq(deseq_ob, test="Wald", fitType="parametric")
deseq_ob_def <- results(deseq_ob, alpha=0.01, contrast=c("region", "Daneborg", "Ittoqqortoormii"))
sigtab_deseq <- deseq_ob_def[which(deseq_ob_def$padj < 0.01), ]
sigtab_deseq_with_tax <- cbind(as(sigtab_deseq, "data.frame"), as(tax_table(taxtable)[row.names(sigtab_deseq), ], "matrix"))
sigtab_deseq_with_tax[order(sigtab_deseq_with_tax$baseMean, decreasing=T), ]
deseq2_ins_fis <- as.data.frame(sigtab_deseq_with_tax)
x = tapply(sigtab_deseq_with_tax$log2FoldChange, sigtab_deseq_with_tax$phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab_deseq_with_tax$phylum = factor(as.character(sigtab_deseq_with_tax$phylum), levels=names(x))
x = tapply(sigtab_deseq_with_tax$log2FoldChange, sigtab_deseq_with_tax$genus, function(x) max(x))
x = sort(x, TRUE)
sigtab_deseq_with_tax$genus = factor(as.character(sigtab_deseq_with_tax$genus), levels=names(x))

colors_alphabetic <- ehi_phylum_colors1 %>%
  right_join(taxonomyclean, by=join_by(phylum == phylum)) %>%
  select(phylum, colors) %>%
	unique() %>%
	dplyr::arrange(phylum)

tax_table1 <- as.data.frame(unique(sigtab_deseq_with_tax$phylum))
colnames(tax_table1)[1] <- "phylum"
tax_color1 <- merge(tax_table1, colors_alphabetic, by="phylum")%>%
	dplyr::arrange(phylum) %>%
	select(colors) %>%
	pull()

sigtab_deseq_with_tax1 <- sigtab_deseq_with_tax %>%	dplyr::arrange(phylum)

```

```{r deseq2_plot, comment="", echo=FALSE, message=FALSE, warning=FALSE}
ggplot(sigtab_deseq_with_tax, aes(x=forcats::fct_rev(reorder(genus, as.integer(factor(phylum)))), y=log2FoldChange, color=phylum)) + 
  geom_point(size=4) + 
  scale_color_manual(values=tax_color1) + 
  geom_hline(yintercept=0) + 
  coord_flip()+
  xlab("Genus") + 
  ylab("log2FoldChange")+
  guides(col=guide_legend("Phylum"))

```